### 2024-10-30 PbD/summary.tex
Added basic layout  for pbd summary.

\documentclass{article}

\usepackage{../.template/summary}

\subject{Privacy by Design}
\semester{Winter 2024}
\author{Leopold Lemmermann}

\usepackage{pdfpages}

\begin{document}\createtitle

\part{Privacy Enhancing Technology}
% \includepdf[pages={?}, nup=?]{slides.pdf}

\section{IPv6 traffic pseudonymisation}

\section{In Practice}

\section{Censorship Resistance}

\section{Criminal Misuse and Data Retention}

\part{Location Privacy}

\end{document}