### 2024-10-30 PbD/summary.tex
Added some basic content for PbD summary.

\documentclass{article}

\usepackage{../.template/summary}

\subject{Privacy by Design}
\semester{Winter 2024}
\author{Leopold Lemmermann}

\usepackage{pdfpages}

\begin{document}\createtitle

\part{Privacy Enhancing Technology}

\section{Confidentiality of Metadata}
Goal of privacy.
- Anonymity: Protection of the identity of the sender or receiver of a message.
- Unobservability: Protection of the communication relation itself.

\section{Attacker model}
- Outsider: eavesdropping on communcation lines, traffic analysis
- Insider: network operators, malicious staff, government
can:
- passively observe all communication links
- actively send messages
- flood or block messages
- operate (all but one) anonymity services
- operate server
cannot:
- break cryptographic systems
- attack user's personal machine
- have unlimited time or resources

\section{Tracking on OSI Application Layer}
Application:
- filtering mechanisms: cookies, active contents
- e2e encryption
- hiding: pseudonyms, credentials

Pseudonyms
- Passive Trustee: trusted third party (TTP) that certifies pseudonyms
- Active Trustee: TTP is involved in communication

\section{Tracking on OSI Network Layer}
- Broadcast with implicit addressing: all recipients receive all messages, local filtering -> protects recipient
- Dummy Traffic: send meaningless messages -> protects sender
- Proxies: forward messages -> protects sender, but proxy can track
- DC network: broadcast, dummy and cryptography -> protects sender
- Blind-message service: unobservable access to replicated databases -> protects database client
- Mix network: cascaded (independent) proxies, dummy traffic, cryptography -> protects communication relation
- Steganography: hidden messages in other messages -> protects content

\section{Traffic Analysis}
Attacks
- Timing: correlation of messages
- Volume: correlation of message sizes
- Flooding: fill up anonymous group batch to separate certain message
- Intersection: correlation of messages in different groups

% \includepdf[pages={14-29}, nup=3x3]{slides.pdf}
\section{IPv6 traffic pseudonymisation}

% \includepdf[pages={?}, nup=?]{slides.pdf}
\section{In Practice}

% \includepdf[pages={?}, nup=?]{slides.pdf}
\section{Censorship Resistance}

% \includepdf[pages={?}, nup=?]{slides.pdf}
\section{Criminal Misuse and Data Retention}

\part{Location Privacy}

\end{document}