### 2024-12-05 SbD/res/des.tbl.tex
Added current state.

\begin{table}[H]
  \centering
  \caption{Ausgewählte DES-Funktionswerte}
  \begin{tabular}{c|c|c|c}
  \textbf{Schlüssel $k$} & \textbf{Eingabe $x/y$ von $E_k/D_k$} & \textbf{$E_k(x)$} & \textbf{$D_k(y)$} \\ \hline
  97a4 3741 8086 0b & 6b3b acd4 6570 6187 & fecd 14e2 8cb0 6a2b & 7285 3f6b e44c cb8e \\ \hline
  97a4 3741 8086 0b & 5de1 bd06 54fa 40ed & 42e0 4462 099e e8ca & 9384 bebf 5b18 793f \\ \hline
  3e45 c68f bb89 27 & 5de1 bd06 54fa 40ed & 4995 e68b b800 3a06 & f575 4696 effc 1bf1 \\ \hline
  e6f4 a450 4898 d7 & 764d e3ae 043b d6b7 & 5f31 4917 4ecc e6b4 & 30db 8c29 073a 182c \\ \hline
  f6bb 747d dc16 15 & ceae a65a ae77 1c72 & b1e0 9414 e068 3622 & 9acb 4aff e84a b1f7 \\ \hline
  3e45 c68f bb89 27 & 764d e3ae 043b d6b7 & fca6 8c2f b81e dc79 & 4bae f15e 7efb 5a8c \\ \hline
  e6f4 a450 4898 d7 & c183 50b5 52b8 f8b1 & 5de1 bd06 54fa 40ed & 6b3b acd4 6570 6187 \\ \hline
  3e45 c68f bb89 27 & 9400 39ca 9301 5507 & 199c 2565 ea8e 2b53 & f742 7dac e2ff 00d2 \\ \hline
  f6bb 747d dc16 15 & c620 9b86 c72a 6b3b & 4e6c efc7 1f5b 3756 & 42e0 4462 099e e8ca \\ \hline
  e6f4 a450 4898 d7 & 42e0 4462 099e e8ca & c3fc 30e9 10db 827e & d0eb 0c0e 1b43 12da \\ \hline
  f6bb 747d dc16 15 & a904 aa4e c934 ecb0 & 42e0 4462 099e e8ca & 5617 0d2a ccdb 9981 \\ \hline
  f6bb 747d dc16 15 & a3c5 5b7b c599 e5af & 4179 a965 d79e 1edd & e927 17a5 b2a9 b559 \\ \hline
  e6f4 a450 4898 d7 & 9400 39ca 9301 5507 & 0669 924e 1cec 70d0 & 7555 ce29 f46b 7e74 \\ \hline
  3e45 c68f bb89 27 & 6b3b acd4 6570 6187 & abe6 c175 4435 742a & 043c 252e bbe0 3d09 \\ \hline
  97a4 3741 8086 0b & c183 50b5 52b8 f8b1 & 9400 39ca 9301 5507 & 764d e3ae 043b d6b7 \\ \hline
  97a4 3741 8086 0b & 42e0 4462 099e e8ca & 8b99 c9de 0e47 22db & 5de1 bd06 54fa 40ed \\
  \end{tabular}
\end{table}